Investigating country-level activity and mobility patterns and their interdependencies on COVID-19 outcomes

### 2020-11-17 docs/manuscript/manuscript.tex
Aggregation Procedure Overview
Updated data dictionary

\usepackage{rotating}
%% Cite Title
%\usepackage[style=numeric,backend=biber,sorting=none,maxcitenames=2,maxbibnames=99,doi=false,isbn=false,url=false,eprint=false]{biblatex}
%\addbibresource{references.bib}

@@

@@

\subsection{Global vector autoregression model with time-varying weights}

% We use a Global Vector AutoRegression (GVAR) modeling framework \cite{dees2007exploring} (Carteni et al
% 2020).
% Given country $i$ with $k_{i}$ endogenous variables $x_{i,t}$, the model is specified as
% \begin{equation}
%   \label{eq:1}
%   x_{i,t} = \sum_{l=1}^{p_{i}}\Phi_{il}x_{i,t-l} + \Lambda_{i0}x_{i,t}^{*} + \sum_{l=1}^{q_{i}} \Lambda_{il}x_{i,t-l}^{*} + \varepsilon_{i,t}
% \end{equation}
% where $i= 1, 2, \ldots, N$. The matrix $\Phi_{il}$ is the matrix of coefficients of size $k_{i}\times k_{i}$, while the matrix $\Lambda_{i0}$ and $\Lambda_{il}$ are $k_{i}\times k_{i}^{*}$ matrices, where $k_{i}^{*}$ is the number of foreign variables.

@@

The basic vector autoregression (VAR) model is given by:
\begin{equation}
  \label{eq:6}
  y_{t} = A_{0}(t) + A(\ell)Y_{t-1} + \varepsilon_{t}
\end{equation}
where $y_{t}$ is a $k\times 1$ vector of endogenous variables for a given unit at time $t\in \{1,\ldots, T\}$,
$A_{0}(t)$ captures the deterministic components and 
$A(\ell)$ is a polynomial in the lag operator and $\varepsilon_{t}\sim iid(0,\Sigma_{\varepsilon})$ are random disturbances.
This framework allows for specification of both dynamic and static interdependencies
among the endogenous variables, as well as cross-sectional heterogeneities (when a panel VAR framework is used).
Country-specific VAR models can also be stacked to yield a global VAR  \cite{martin2017weighting,dees2007exploring,covid19b}, allowing for the inclusion of weakly exogenous (``foreign'') variables (such as border closures) using weights (such as air travel flows) from each of the countries.

>>>>

\usepackage{rotating}
\usepackage{amsmath, amssymb}
%% Cite Title
%\usepackage[style=numeric,backend=biber,sorting=none,maxcitenames=2,maxbibnames=99,doi=false,isbn=false,url=false,eprint=false]{biblatex}
%\addbibresource{references.bib}

@@

\subsection{Aggregation Procedure}
Four of the data sets had to be aggregated in order to standardize the GVAR model inputs.

  \begin{enumerate}
  \item New Cases and Deaths.
   Both new cases and death statistics are provided by JHU and follow the same format. First we sum up province data into a country total by checking each row if the 'province/state' field is NA.
   If that is not the case, we add all rows for the entire series. Then we incorporate the 'ISO' based on 3-digit alpha codes. Finally, we apply the melt function to the newly created DataFrame
   and use 'pivot table', with iso and activity set as index.
  \item Google Activity.
   We retain a subset of the original data containing national stats without the subregional breakdown. Afterwards, we rename activity types for conventional purposes (i.e. retail and recreation  as retail). Again, we apply melt and 'pivot table' with ISO and activity set as index. 2-digit country codes are replaced with the 3-digit ISO convention.
  \item Government Interventions.
  We create a new DataFrame with 'MultiIndex' for Category and ISO. We ensure the ISO code conforms to the 3-digit convention.
  \end{enumerate}

@@

\subsection{Global vector autoregression model with time-varying weights}
The global VAR model \cite{dees2007exploring} is given by

\begin{equation}
  \label{eq:1}
  x_{i,t} = \sum_{l=1}^{p_{i}}\Phi_{il}x_{i,t-l} + \Lambda_{i0}x_{i,t}^{*} + \sum_{l=1}^{q_{i}}\Lambda_{il}x_{i,t-l}^{*} + \varepsilon_{i,t}
\end{equation}
where $x_{i,t}$ is a vector of $k_{i}\times 1$ endogenous variables and $i = 1, 2, \ldots, N$ is the number of buildings
in the system.  The matrix $\Phi_{il}$ has dimensions $k_{i}\times k_{i}$, while the matrices $\Lambda_{io}$ and
$\Lambda_{i1}$ both have the dimensions $k_{i}\times k_{i}^{*}$.
Each country $i$ has $k_{i}$ endogenous variables and $k_{i}^{*}$ foreign (weakly exogenous) variables.
$\Phi_{il}$ is a coefficient matrix of the cointegrated
endogenous variables. $\Lambda_{i0}$ is a coefficient matrix for the contemporanoues weakly exogenous variable
$x_{i,t}^{*}$. $\Lambda_{il}$ is the coefficient matrix of the $q_{i}$ lagged foreign variables $x_{i,t-l}^{*}$, where
\begin{equation}
  \label{eq:2}
  x_{i,t}^{*} = \ol{W}_{i}^{'}x_{t}.
\end{equation}
The matrix $\ol{W}_{i}^{'}$ is the $k^{*} \times N$ matrix of building-specific weights.
Stacking the domestic and foreign variables, we obtain
\begin{equation}
  \label{eq:3}
  z_{i,t} =
  \begin{bmatrix}
    x_{i,t}^{'}, x_{i,t}^{*'}
  \end{bmatrix}'  = W_{i}x_{t}
\end{equation}
where $W_{i} = E_{i}\ol{W}_{i}$ and $E_{i}$ is a selection matrix We
can then compactly write the model as
\begin{equation}
  \label{eq:4}
  G_{0}x_{t} = \sum_{l=1}^{p}G_{l}x_{t-l} + \varepsilon_{t}
\end{equation}
where
\begin{equation}
  \label{eq:5}
  G_{l} =
  \begin{bmatrix}
    A_{1,l}W_{1}\\ A_{2,l}W_{2}\\ \vdots \\A_{N,l}W_{N}
  \end{bmatrix}
  % =
  % \begin{bmatrix}
  %   \Phi_{1l} & \Lambda_{2l} \\ \vdots  & \vdots \\     \Phi_{Nl} & \Lambda_{Nl} 
  % \end{bmatrix}
\end{equation}
and $A_{i,l} = [\Phi_{1l}  \Lambda_{2l}]$.

The global model can then be compactly written as
\begin{equation}
  \label{eq:6}
  x_{t} = \sum_{l=1}^{p}G_{0}^{-1}G_{l }x_{t-l} + G_{0}^{-1}u_{t}
\end{equation}
where $x_{t}$ $x_{t-l}$ and $u_{t}$ are stacked vectors of length $kN$.

% Given country $i$ with $k_{i}$ endogenous variables $x_{i,t}$, the model is specified as
% \begin{equation}
%   \label{eq:1}
%   x_{i,t} = \sum_{l=1}^{p_{i}}\Phi_{il}x_{i,t-l} + \Lambda_{i0}x_{i,t}^{*} + \sum_{l=1}^{q_{i}} \Lambda_{il}x_{i,t-l}^{*} + \varepsilon_{i,t}
% \end{equation}
% where $i= 1, 2, \ldots, N$. The matrix $\Phi_{il}$ is the matrix of coefficients of size $k_{i}\times k_{i}$, while the matrix $\Lambda_{i0}$ and $\Lambda_{il}$ are $k_{i}\times k_{i}^{*}$ matrices, where $k_{i}^{*}$ is the number of foreign variables.

@@

% The basic vector autoregression (VAR) model is given by:
% \begin{equation}
%   \label{eq:6}
%   y_{t} = A_{0}(t) + A(\ell)Y_{t-1} + \varepsilon_{t}
% \end{equation}
% where $y_{t}$ is a $k\times 1$ vector of endogenous variables for a given unit at time $t\in \{1,\ldots, T\}$,
% $A_{0}(t)$ captures the deterministic components and 
% $A(\ell)$ is a polynomial in the lag operator and $\varepsilon_{t}\sim iid(0,\Sigma_{\varepsilon})$ are random disturbances.
% This framework allows for specification of both dynamic and static interdependencies
% among the endogenous variables, as well as cross-sectional heterogeneities (when a panel VAR framework is used).
% Country-specific VAR models can also be stacked to yield a global VAR  \cite{martin2017weighting,dees2007exploring,covid19b}, allowing for the inclusion of weakly exogenous (``foreign'') variables (such as border closures) using weights (such as air travel flows) from each of the countries.
